Diverse Genotype-by-Weather Interactions in Switchgrass

### 2024-06-12 R/Analysis_v1.2_mash_using_greedy_covar.qmd
revision2: generate revised figures 1-4 and SI figures 1-4

@@

@@

save_plot(filename = "../manuscript/revision2/Figure_2_recolor_covariance_matrices_and_posterior_weights.svg", plot = last_plot(), base_height = 5.5, base_width = 8)
```

>>>>

NB: Figure 1 is in Plots_v0.3_manuscript_figures.Rmd

@@

Build Figure 2 with covariance matrices as first two rows over covar barplots

ggdraw() +
  theme_oeco + 
  draw_plot(pa01, x = 0, y = 0.66, width = 0.1, height = 0.18) +
  draw_plot(pa02, x = 0.2, y = 0.66, width = 0.1, height = 0.18) +
  #draw_plot(pa03, x = -0.02, y = 0.6, width = 0.1, height = 0.18) +
  #draw_plot(pa04, x = 0.1, y = 0.6, width = 0.1, height = 0.18) +
  draw_plot(pa05, x = 0.4, y = 0.66, width = 0.1, height = 0.18) +
  draw_plot(pa06, x = 0.6, y = 0.66, width = 0.1, height = 0.18) +
  draw_plot(pa07, x = 0.8, y = 0.66, width = 0.1, height = 0.18) +
  #draw_plot(pa08, x = 0.2, y = 0, width = 0.1, height = 0.18) +
  draw_plot(pa09, x = 0, y = 0.83, width = 0.1, height = 0.18) +
  draw_plot(pa10, x = 0.2, y = 0.83, width = 0.1, height = 0.18) +
  draw_plot(pa11, x = 0.4, y = 0.83, width = 0.1, height = 0.18) +
  draw_plot(pa12, x = 0.6, y = 0.83, width = 0.1, height = 0.18) +
  draw_plot(pa13, x = 0.8, y = 0.83, width = 0.1, height = 0.18) +
  draw_plot(f2a, x = 0, y = 0.33, width = 0.66, height = 0.38) +
  draw_plot(legend, x = 0.1, y = 0.77, width = 0.1, height = 0.2) +
  draw_plot(f2b, x = 0.66, y = 0.33, width = 0.33, height = 0.38) +
  draw_plot(f2c, x = 0, y = 0, width = 0.66, height = 0.33) +
  draw_plot(f2d, x = 0.66, y = 0, width = 0.33, height = 0.33)
```

@@

save_plot(filename = "../manuscript/revision2/Figure_2_recolor_resize_covariance_matrices_and_posterior_weights.svg", plot = last_plot(), base_height = 7.5, base_width = 6.5)

### 2024-06-12 R/Plots_v0.3_manuscript_figures.Rmd
revision2: generate revised figures 1-4 and SI figures 1-4

Add any libraries/scripts that need to be loaded for these analyses here.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(viridis)
library(lubridate)
library(switchgrassGWAS)
library(GGally)
library(maps)
library(ggmap)
usa <- map_data("usa")
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
```

# Load data
Add any data that needs to be loaded for these analyses here.
```{r}
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
phe_gr_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
phe_fl_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
phenotypes <- readRDS("~/Github/pvdiv-phenology-gxe/data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
h2gr <- read_csv(file.path(workingdir, "analysis", "heritability", "kinship_subpop", "h2_df_GR50_related_functions_of_weather_v1.csv"))
h2fl <- read_csv(file.path(workingdir, "analysis", "heritability", "kinship_subpop", "h2_df_FL50_related_functions_of_weather_v1.csv"))
h2gr_gxe <- read_csv(file.path(workingdir, "analysis", "heritability", "kinship_subpop", "gxe_and_h2_df_GR50_related_functions_of_weather_v2.csv"))
h2fl_gxe <- read_csv(file.path(workingdir, "analysis", "heritability", "kinship_subpop", "gxe_and_h2_df_FL50_related_functions_of_weather_v1.csv"))

h2gr <- h2gr %>%
  full_join(h2gr_gxe)
h2gr <- h2gr_gxe
h2fl <- h2fl %>%
  full_join(h2fl_gxe)
```
as well as the code for any new analyses that I'm running to do new writeup for the manuscript v0.3.

@@

ggplot() + 
  geom_polygon(data = usa, aes(x=long, y = lat, group = group), 
               fill = "white", color = "black") +
  coord_quickmap(xlim = c(-105, -66), ylim = c(25, 50)) +
  geom_jitter(data = metaplot, aes(x = LONGITUDE, y = LATITUDE, 
                                   color = SUBPOP, shape = ECOTYPE_NNET),
              width = 0.45, height = 0.45) + 
  geom_point(data = siteplot, aes(x = Longitude, y = Latitude), shape = 21, size = 4, stroke = 1.5, fill = "white") +
  geom_point(data = siteplot, aes(x = Longitude, y = Latitude), shape = 4, size = 3, stroke = 1.5) +
  geom_label_repel(data = siteplot, aes(x = Longitude, y = Latitude, label = manu_site), alpha = 0.7) +
  labs(x = "Longitude", y = "Latitude") + 
  theme(legend.position = "right") +
  scale_color_manual(values = c("#F47F72", "#442C83")) +
  scale_shape_manual(values = c(0, 6, 1, 4)) 
  #scale_fill_manual(values = c("#280B54FF","#FAC127FF", "#D44842FF", "white")) +
  #scale_color_manual(values = c(rep("black", 10), "#280B54FF","#FAC127FF", "#D44842FF", "white", "#280B54FF","#FAC127FF", "#D44842FF", "white"), guide = FALSE) +
  #scale_alpha_manual(values = c(0.7, 0.7, 0.7, 0.7), guide = FALSE)
save_plot(filename = "../manuscript/plots/Map_subpops_and_Ecotypes.png", plot = last_plot(), base_height = 3)
?geom_point
```

@@

```{r}
pand <- cor_fl %>%
  filter(SUBPOP %in% c("Gulf")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = FALSE, layout.exp = 0.2, hjust = 0.95) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  theme(legend.position = "none")
pane <- cor_fl %>%
  filter(SUBPOP %in% c("Midwest")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = FALSE, layout.exp = 0.2, hjust = 0.95) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  theme(legend.position = "none")
panf <- cor_fl %>%
  filter(SUBPOP %in% c("Gulf", "Midwest")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = FALSE, layout.exp = 0.2, hjust = 0.95) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  theme(legend.position = "none")
pana <- cor_gr %>%
  filter(SUBPOP %in% c("Gulf")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = FALSE, layout.exp = 0.2, hjust = 0.95) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  theme(legend.position = c(0.4, 0.85))
panb <- cor_gr %>%
  filter(SUBPOP %in% c("Midwest")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = FALSE, layout.exp = 0.2, hjust = 0.95) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  theme(legend.position = "none")
panc <- cor_gr %>%
  filter(SUBPOP %in% c("Gulf", "Midwest")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = FALSE, layout.exp = 0.2, hjust = 0.95) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  theme(legend.position = "none")

pheno_corr <- plot_grid(pana, panb, panc, pand, pane, panf, nrow = 2)
save_plot(filename = "../manuscript/plots/Phenotypic_correlations_GR50_FL50_2.svg", plot = pheno_corr, base_width = 6.5)
```

>>>>

Add any libraries/scripts that need to be loaded for these analyses here.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(cowplot)
library(viridis)
library(lubridate)
library(switchgrassGWAS)
library(GGally)
library(maps)
library(ggmap)
usa <- map_data("usa")
# source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
```

```{r}
theme_oeco <- theme_classic() +
  theme(axis.title = element_text(size = 10), 
        axis.text = element_text(size = 10), 
        axis.line.x = element_line(linewidth = 0.35, colour = 'grey50'),
        axis.line.y = element_line(linewidth = 0.35, colour = 'grey50'),
        axis.ticks = element_line(linewidth = 0.25, colour = 'grey50'), 
        legend.justification = c(1, 0.75), 
        legend.position = c(1, 0.9), 
        legend.key.size = unit(0.35, 'cm'),
        legend.title = element_blank(), 
        legend.text = element_text(size = 9),
        legend.background = element_blank(),
        plot.subtitle = element_text(size = 10, vjust = 0), #plot.margin = unit(c(0.35, 0, 0.25, 0), 'cm'),
        strip.background = element_blank(), 
        strip.text = element_text(hjust = 0.5, size = 10 ,vjust = 0), 
        strip.placement = 'outside', panel.spacing.x = unit(-0.5, 'cm'))
theme_set(theme_oeco)
```

# Load data
Add any data that needs to be loaded for these analyses here.
```{r}
# workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
metadata <- read_rds(here("data", "metadata.rds"))
sites <- read_rds(here("data", "sites.rds"))
subpops <- read_rds(here("data", "subpop_color_coding.rds"))
phe_gr_df <- read_rds(here("data",
                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
phe_fl_df <- read_rds(here("data",
                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
phenotypes <- readRDS(here("data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds"))
h2gr <- read_csv(here("analysis", "heritability", "kinship_subpop", "h2_df_GR50_related_functions_of_weather_v1.csv"))
h2fl <- read_csv(here("analysis", "heritability", "kinship_subpop", "h2_df_FL50_related_functions_of_weather_v1.csv"))
h2gr_gxe <- read_csv(here("analysis", "heritability", "kinship_subpop", "gxe_and_h2_df_GR50_related_functions_of_weather_v2.csv"))
h2fl_gxe <- read_csv(here("analysis", "heritability", "kinship_subpop", "gxe_and_h2_df_FL50_related_functions_of_weather_v1.csv"))

#h2gr <- h2gr %>%
#  full_join(h2gr_gxe)
h2gr <- h2gr_gxe
h2fl <- h2fl %>%
  full_join(h2fl_gxe)
```
as well as the code for any new analyses that I'm running to do new writeup for the manuscript v0.3.

@@

ggplot() + 
  geom_polygon(data = usa, aes(x=long, y = lat, group = group), 
               fill = "white", color = "black") +
  coord_quickmap(xlim = c(-105, -66), ylim = c(25, 50)) +
  geom_jitter(data = metaplot, aes(x = LONGITUDE, y = LATITUDE, 
                                   color = SUBPOP, shape = ECOTYPE_NNET),
              width = 0.45, height = 0.45) + 
  geom_point(data = siteplot, aes(x = Longitude, y = Latitude), shape = 21, size = 4, stroke = 1.5, fill = "white") +
  geom_point(data = siteplot, aes(x = Longitude, y = Latitude), shape = 4, size = 3, stroke = 1.5) +
  geom_label_repel(data = siteplot, aes(x = Longitude, y = Latitude, label = manu_site), alpha = 0.7) +
  labs(x = "Longitude", y = "Latitude") + 
  theme(legend.position = "right") +
  scale_color_manual(values = c("#F47F72", "#442C83")) +
  scale_shape_manual(values = c(0, 6, 1, 4)) 
  #scale_fill_manual(values = c("#280B54FF","#FAC127FF", "#D44842FF", "white")) +
  #scale_color_manual(values = c(rep("black", 10), "#280B54FF","#FAC127FF", "#D44842FF", "white", "#280B54FF","#FAC127FF", "#D44842FF", "white"), guide = FALSE) +
  #scale_alpha_manual(values = c(0.7, 0.7, 0.7, 0.7), guide = FALSE)
save_plot(filename = here("manuscript", "revision2", "intermediate_files", "Map_subpops_and_Ecotypes.png"), plot = last_plot(), base_height = 3)
?geom_point
```

@@

```{r}
pand <- cor_fl %>%
  filter(SUBPOP %in% c("Gulf")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = FALSE, layout.exp = 0.2, hjust = 0.95) +
  scale_fill_gradientn(colors = c("#30123BFF", "#4777EFFF", "#1BD0D5FF",
                                  "white", "#FE9B2DFF", "#DB3A07FF",
                                  "#7A0403FF"), limits = c(-1,1)) +
  theme(legend.position = "none")
pane <- cor_fl %>%
  filter(SUBPOP %in% c("Midwest")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = FALSE, layout.exp = 0.2, hjust = 0.95) +
  scale_fill_gradientn(colors = c("#30123BFF", "#4777EFFF", "#1BD0D5FF",
                                  "white", "#FE9B2DFF", "#DB3A07FF",
                                  "#7A0403FF"), limits = c(-1,1)) +
  theme(legend.position = "none")
panf <- cor_fl %>%
  filter(SUBPOP %in% c("Gulf", "Midwest")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = FALSE, layout.exp = 0.2, hjust = 0.95) +
  scale_fill_gradientn(colors = c("#30123BFF", "#4777EFFF", "#1BD0D5FF",
                                  "white", "#FE9B2DFF", "#DB3A07FF",
                                  "#7A0403FF"), limits = c(-1,1)) +
  theme(legend.position = "none")
pana <- cor_gr %>%
  filter(SUBPOP %in% c("Gulf")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = FALSE, layout.exp = 0.2, hjust = 0.95) +
  scale_fill_gradientn(colors = c("#30123BFF", "#4777EFFF", "#1BD0D5FF",
                                  "white", "#FE9B2DFF", "#DB3A07FF",
                                  "#7A0403FF"), limits = c(-1,1)) +
  theme(legend.position = c(0.4, 0.85))
panb <- cor_gr %>%
  filter(SUBPOP %in% c("Midwest")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = FALSE, layout.exp = 0.2, hjust = 0.95) +
  scale_fill_gradientn(colors = c("#30123BFF", "#4777EFFF", "#1BD0D5FF",
                                  "white", "#FE9B2DFF", "#DB3A07FF",
                                  "#7A0403FF"), limits = c(-1,1)) +
  theme(legend.position = "none")
panc <- cor_gr %>%
  filter(SUBPOP %in% c("Gulf", "Midwest")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = FALSE, layout.exp = 0.2, hjust = 0.95) +
  scale_fill_gradientn(colors = c("#30123BFF", "#4777EFFF", "#1BD0D5FF",
                                  "white", "#FE9B2DFF", "#DB3A07FF",
                                  "#7A0403FF"), limits = c(-1,1)) +
  theme(legend.position = "none")

pheno_corr <- plot_grid(panc, panf, panb, pane, pana, pand, nrow = 3)
save_plot(filename = here("manuscript", "revision2", "intermediate_files", "Phenotypic_correlations_GR50_FL50_2.svg"), plot = pheno_corr, base_width = 3.5, base_height = 4.5)
```

### 2024-06-12 R/SI_Appendix_Datasets.Rmd
revision2: generate revised figures 1-4 and SI figures 1-4

Edited 2024-05-17 for revision2.

@@

Figure 3 is North x Texas GxE effects. Figure S1 is garden x garden GxE effects of individual loci.

@@

## Figure S2

@@

We conducted quantitative trait loci (QTL) mapping of flowering as functions of five environmental cues that we also used as covariance matrices in mash, and identified eight QTL for flowering date, eight QTL for flowering as a function of day length change two days prior, one QTL for the start of vegetative growth, and two QTL for vegetative growth as a function of daylength change one day prior, all of which showed QTL by environment interactions (SI Appendix, Figure S2).

>>>>

Edited 2024-05-17+ for revision2.

@@

h^2 analyses that was originally figure 1C. Try moving to the supplement.

## Figure S2 & S3

Figure 3 is North x Texas GxE effects. Figure S2 is garden x garden GxE effects of individual loci for green-up; Figure S3 is for flowering date.

@@

## Figure S4

@@

We conducted quantitative trait loci (QTL) mapping of flowering as functions of five environmental cues that we also used as covariance matrices in mash, and identified eight QTL for flowering date, eight QTL for flowering as a function of day length change two days prior, one QTL for the start of vegetative growth, and two QTL for vegetative growth as a function of daylength change one day prior, all of which showed QTL by environment interactions (SI Appendix, Figure S4).

```{r}
library(qtl)
library(qtl2)
library(qtl2convert)
```

```{r prep QTL_Eff df}
###QTL Effect plot
###load the original cross and reduce the markers from 4700 to 1185
#load('../phenology-fourway/5-QTLmappingForAlice/sexAveraged.cross.rda', verbose = T)
workingdir <- here("phenology-fourway/5-QTLmappingForAlice/")
load(file.path(workingdir, 'cross_reduced_marker_FL50_phenotypes_Alice_7_SITES.RData'))
map  =  lapply(pull.map(cross), function(x) x[1,])
# reduced_markers  =  unlist(
#   lapply(
#     reduce_markers(
#       lapply(pull.map(cross), function(x)
#         x[1,]), min_distance = 1), names))
#
# mars2drop  =  markernames(cross)[!markernames(cross) %in% reduced_markers]
# cross  =  drop.markers(cross, mars2drop)
cross = calc.genoprob(cross, step=1, error.prob = 1e-3,map.function = "kosambi",stepwidth = "max")
######### Converting probalibity from qtl1 to qtl2
probs  =  probs_qtl_to_qtl2(cross)$probs
##calculate the kinship matrix for single site
kinship_loco_Single = calc_kinship(probs,use_allele_probs = F,type = "loco")

###extract the genotype ID from the original cross file, so we can align the phenology file with genotype ID
ID <- data.frame(ID=cross$pheno$id)
ID <- ID %>% 
  mutate_at(vars('ID'), list(as.character)) %>%
  mutate_at(vars('ID'), list(as.numeric))

######get the phenotype at each site
phenology = read.csv('../phenology-fourway/5-QTLmappingForAlice/FL50_phenotypes_Alice.csv', header = TRUE)
colnames(phenology)[1:2] = c('Covariate','ID')
str(phenology) ###make sure your phenology traits are numeric, if not, change them into numeric
phenos <- phenology %>% 
  mutate_at(vars('Covariate'), list(as.character)) %>%
  mutate(ID = as.numeric(ID)) %>% 
  group_by(Covariate,ID) %>%
  summarize_all(mean, na.rm=T) %>% 
  ungroup() %>% 
  as.data.frame()
```

```{r setup QTL effects for all traits}
###QTL effect for each trait
EffectMatrix = vector('list', length=0)
t=0
for (envi in unique(phenos$Covariate)){
    t=t+1
    pheno = phenos %>% 
      filter(Covariate==envi) %>% 
      dplyr::select(-Covariate) %>%
      select(ID, FL50) %>% 
      right_join(ID) %>%  column_to_rownames('ID') %>% as.matrix()

    ##ran effect scan for each chromosome
    eff_out = vector('list',length = 0)
    se_out = vector('list',length = 0)
    add_contrasts=cbind(mu=c(1,1,1,1), add1=c(-1, 1, -1, 1),add2=c(-1, -1, 1, 1), dom=c(-1, 1, 1, -1))

    z = 0
    # chr = "1K"
    for (chr in chrnames(cross)){
      print(c(chr, envi,t))
      z= z+1
      eff = scan1coef(probs[,chr],pheno,
                      kinship = kinship_loco_Single[chr], 
                      se=T, contrasts = add_contrasts)
      se = attr(eff, 'SE')
      eff_out[[z]] =  eff %>% as.data.frame() %>% rownames_to_column('MARKER')
      se_out[[z]] =  se %>% as.data.frame() %>% rownames_to_column('MARKER')
    }
    eff_out = bind_rows(eff_out)
    se_out = bind_rows(se_out)

    se_out2 = se_out %>% dplyr::select(-mu) %>% gather('CROSS','SE',-MARKER)
    eff_out2 = eff_out %>% dplyr::select(-mu) %>% gather('CROSS','EFF',-MARKER) %>% mutate(SE=se_out2$SE, Covariate=envi)

    EffectMatrix[[t]] = eff_out2
  }

EffectMatrix2 = bind_rows(EffectMatrix)

###unlist nmap and merge with Effect file
nmap = data.frame(pos = unlist(map))
nnmap = nmap %>% rownames_to_column('MARKER') %>% rowwise() %>%
  mutate(chr = str_sub(MARKER, 1,2), MARKER=str_sub(MARKER, 4,nchar(MARKER)))%>%
  ungroup() %>% mutate(pos=round(pos,4))

flank_marker_full = read.csv('../phenology-fourway/5-QTLmappingForAlice/QTLsWithFlankMarkers_FL50_phenotypes_Alice_7_SITES_Full_model.csv')
flank_marker_full = flank_marker_full %>% mutate_at(vars(chr),list(as.character))

for (i in 1:nrow(flank_marker_full)){
  tmp = nnmap %>% filter(chr==flank_marker_full$chr[i] & pos==round(flank_marker_full$pos[i],4))
  flank_marker_full$MARKER[i] = tmp$MARKER
}

QTL_full_model = flank_marker_full %>% dplyr::rename(trait=lodcolumn)

```{r ggplots of QTL effects}
QTLdf <- QTL_full_model %>% left_join(EffectMatrix2)%>%
  mutate(CROSS = str_replace(CROSS,'add2','C x D')) %>%
  mutate(CROSS = str_replace(CROSS,'add1','A x B'),
         Cross = case_when(CROSS == "A x B" ~ "AP13 x DAC",
                           CROSS == "C x D" ~ "VS16 x WBC",
                           CROSS == "dom" ~ "Dominance",
                           TRUE ~ "Other"),
         Trait = case_when(trait == "dyln_change_sec" ~ "Daylength change (s), on Green-up Date",
                           trait == "dyln_change_sec_2d_prior" ~ "Daylength change (s), 2d prior to Flowering Date",
                           trait == "FL50" ~ "Flowering Date",
                           trait == "GR50" ~ "Green-up Date",
                           TRUE ~ "Other")
         ) %>%
  separate(MARKER, into = c("Chr", "Pos"), sep = "_", convert = TRUE, 
           remove = FALSE) %>%
  mutate(Pos_plot = paste0(Chr, "\n", round(Pos, 1), " Mb"))

QTLdf$Covariate = factor(QTLdf$Covariate, levels = c("TX1", "TX2", "OK", "MO", 
                                                   "NE", "MI", "SD"))

QTLdf %>% 
  #filter(Cross != "Dominance") %>% # trait == 'FL50' &
  ggplot(aes(x = Covariate, y = EFF, group = Cross)) +
  geom_point(size = 2) + facet_grid(Cross + trait ~ Pos_plot) +
  geom_hline(yintercept = 0, linetype=2) + 
  geom_errorbar(aes(ymin=EFF-SE, ymax=EFF+SE), size=0.6, width=0.5) +
  xlab("") +
  ylab('QTL Effect') + 
  coord_flip() + 
  theme_bw()+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        text=element_text(size=12, face="bold"),
        axis.title=element_text(size=12,face="bold"),
        axis.text.x = element_text(size = 10, angle = 45, hjust = 1, vjust = 1))
# shared QTL for more than one trait differ in LOD score, but not meaningfully in effect patterns across gardens. Useful to know.

QTLplot <- QTLdf %>%
  arrange(desc(lod)) %>%
  group_by(Chr, Pos, Covariate, Cross) %>%
  slice(1) %>% 
  filter(Cross != "Dominance") # trait == 'FL50' &

QTLplot %>%
  ggplot(aes(x = Covariate, y = EFF, group = Cross)) +
  geom_col(size = 1, aes(fill = Trait)) + 
  facet_grid(Cross ~ Pos_plot) +
  geom_hline(yintercept = 0, linetype=2) + 
  geom_errorbar(aes(ymin=EFF-SE, ymax=EFF+SE, color = Trait), size=0.6, width=0.5) +
  xlab("") + scale_y_continuous(limits = c(-2.5, 4), breaks = c(-2, 0, 2)) +
  ylab('QTL Effect') + 
  coord_flip() + 
  switchgrassGWAS::theme_oeco + 
  scale_color_manual(values = c("#CC0066", "#FFCC00", "darkgrey", "darkgreen")) +
  scale_fill_manual(values = c("#CC0066", "#FFCC00", "darkgrey", "darkgreen")) +
  theme(legend.position = "top",
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        text = element_text(size = 10, face="bold"),
        strip.text = element_text(size = 10, face="bold"),
        axis.title = element_text(size = 10,face="bold"),
        axis.text.x = element_text(size = 10, angle = 45, hjust = 1, vjust = 1),
        )# +
    #scale_y_continuous(expand = c(0.2, 0.2)) 
save_plot(filename = here("manuscript/revision2/intermediate_files/ Figure_S4_QTL_effects_by_trait.svg"), base_height = 4.5, base_width = 9, plot = last_plot())

QTLdf %>%
  filter(trait == 'dyln_change_sec_2d_prior') 
QTLdf %>% 
  filter(trait == 'FL50')
QTLdf %>% 
  filter(trait == 'dyln_change_sec')

unique(QTLdf$trait)
```

### 2024-06-12 manuscript/revision2/SI_Appendix/SI_Appendix.qmd
revision2: generate revised figures 1-4 and SI figures 1-4

\noindent This PNAS template for Supporting Information (SI) may be used to organize your supporting material. The template is intended to provide a clearly organized PDF file that will ensure readers can easily navigate to sections or specific figures and tables. Movie files or large datasets can be presented as separate files. Further information is available in the [PNAS Author Center](https://www.pnas.org/authors/submitting-your-manuscript).

## Using the template {.unnumbered}

Specify the title, author list, and corresponding authors with the same YAML specification as in the [corresponding PNAS Quarto template](https://github.com/christopherkenny/pnas). The cover page will be automatically generated with the relevant description of the SI.

Figures should be placed on separate pages with legends set immediately below each figure. Table titles should be set immediately above each table. Note that tables extending beyond the width of the page can be included in the PDF or provided as separate dataset files. Oversized/nonstandard page sizes are accepted as part of your SI Appendix file.

## Submitting SI {.unnumbered}

Save your completed SI file as a PDF for submission. Further submission instructions are available [here](https://www.pnas.org/page/authors/submission#preparation).

@@

@@

where $\mu$ is the population mean, QTL is the marker genetic effect, E is the environmental effects (here, common garden), QTL\*E is the interaction between marker genetic and environmental effects, kinship corresponds to the background polygenic variation, and e is the error term. The genome scan was accomplished with the ‘scan1’ function. The statistical significance of the genome scan was established by performing a stratified (i.e., stratifying on common garden) permutation test (n=1000) using ‘scan1perm’ function. The estimated QTL effect was obtained using ‘scan1coef’ function in R/qtl2.

Figures should be cross-referenced like @fig-qtl-by-env. Each figure should be on its own page. You can control this by placing a pagebreak shortcode, with `{{< pagebreak >}}`.

This can also be done with LaTeX's `\newpage` command. You may find `\FloatBarrier` useful if your figures or tables are running off to the wrong pages.

@@

![Figure](figs/frog.eps){#fig-garden-by-garden-gxe}

@@

![Figure](figs/frog.eps){#fig-qtl-by-env}

>>>>

@@

At this stage, we reviewed the univariate GWAS results, both Manhattan and QQ-plots, for population structure issues that were not resolved using the $k$ that minimized $\lambda_{GC}$. These issues were apparent in the data in two ways: (i) univariate GWAS with many significant SNPs with extremely similar *p*-values, usually indicative of genetic effects that, by chance or, more likely, because of confounding effects of population structure, were identical across many SNPs and thus not resolvable to a single genomic location; (ii) univariate GWAS with many effect estimates and standard errors of zero, indicating an inability to estimate effects given the constraints of small population size & population structure. We only noticed these issues in our smallest subpopulation, the Midwest subpopulation, in the green-up date trait at the SD garden, and in the flowering date trait at the OK and NE gardens. Due to these issues, we dropped the SD garden from subsequent green-up analyses for the Midwest subpopulation, and we dropped the OK and NE gardens from subsequent Midwest flowering date analyses.

@@

where $\mu$ is the population mean, QTL is the marker genetic effect, E is the environmental effects (here, common garden), QTL\*E is the interaction between marker genetic and environmental effects, kinship corresponds to the background polygenic variation, and e is the error term. The genome scan was accomplished with the ‘scan1’ function. The statistical significance of the genome scan was established by performing a stratified (i.e., stratifying on common garden) permutation test (n=1000) using ‘scan1perm’ function. The estimated QTL effects (@fig-qtl-by-env) were obtained using ‘scan1coef’ function in R/qtl2.

@@

![Narrow sense heritability of green-up date and flowering date within single common gardens (purple) and across all eight common gardens (green), within and between two genetic subpopulations.](images/Figure_S1_h2.png){#fig-garden-by-garden-gxe}

@@

![Types of GxE present between of pairs of jointly re-estimated effects in eight common gardens, for effects with lfsr \< 0.05 at both gardens in each pair. Effect patterns: Sign: Effects that differ in sign at these pairs of gardens (p \< 0.05, lfsr). Magnitude: Effects identical in sign (p \< 0.05, lfsr) that differ in magnitude by a factor of \>0.4. Not Distinguishable: Effects not distinguishable by magnitude nor sign of the effect, with no measurable GxE. b) The number of effects with each GxE type for the onset of vegetative growth (green-up date) within and between two genetic subpopulations.](images/S2_Greenup.png)

![Types of GxE present between of pairs of jointly re-estimated effects in eight common gardens, for effects with lfsr \< 0.05 at both gardens in each pair. Effect patterns: Sign: Effects that differ in sign at these pairs of gardens (p \< 0.05, lfsr). Magnitude: Effects identical in sign (p \< 0.05, lfsr) that differ in magnitude by a factor of \>0.4. Not Distinguishable: Effects not distinguishable by magnitude nor sign of the effect, with no measurable GxE. b) The number of effects with each GxE type for the onset of reproductive growth (green-up date) within and between two genetic subpopulations.](images/S3_Flowering.png)

![All QTL identified in the outbred mapping population displayed GxE, in that all QTL had effects that differed in at least two gardens. Gardens are on the y-axis in latitudinal order; the effects of the significant QTL in Figure 4 are displayed as differences between allelic effects for the two sets of grandparent lines in the outbred mapping panel. ](images/Figure_S4_QTL_effects_by_trait.png){#fig-qtl-by-env}
